Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2020-10-30 paper/ifo-complexAudio.tex
adding new version of big filtering plot to include background noise and removing what was fig 9

\begin{figure*}
	\includegraphics[width=.7\textwidth]{figures/setup_pic2.pdf}
\caption{Photograph of the optical microphone. The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right. The interferometer is laid out as in Figure~\ref{fig:ifo_schematic_podo}. In the main image, the photodiode is placed behind the cloth screen. The inset shows a face-on view of the photodiode with the covering screen removed. See Appendix~\ref{app:circuit_diagram} for a circuit diagram and detailed photograph of the circuit.}
	\label{fig:setup_pic2}
\end{figure*}

@@

\begin{figure*}
\begin{center}
% trim order is <left> <lower> <right> <upper>
\includegraphics[width=\textwidth,trim={1cm 8.5cm 1cm 8cm},clip]{figures/freqSpectrumOriginalNoiseSubtracted.pdf}
\end{center}
\caption{\label{fig:noiseSubtract}
Frequency spectrum of a $1\,{\rm s}$ signal before the use of filters. 
The first panel shows the original frequency spectrum of the human voice recording (saying ``A cathode ...''). 
The second panel shows the spectrum recorded after the signal passes through the optical microphone (sampled at $16\,000\,{\rm Hz}$).
The third panel shows a noise-only recording from the optical microphone when no audio signal is being played. 
The fourth panel shows the result of subtracting the noise-only spectrum in the third panel from the recording in the second panel. 
We see no obvious improvement from this method. 
}
\end{figure*}

The Fourier spectrum of the source signal is shown in the first panel of Fig.~\ref{fig:noiseSubtract}. The second panel shows the spectrum of the raw optical microphone recording of the signal. The third panel of Fig.~\ref{fig:noiseSubtract} shows a background spectrum of the optical microphone, as shown previously in the top panel Fig.~\ref{fig:psd_noise}.

Given that we have access to the background noise spectrum, an intuitive way to remove noise is via subtracting the noise spectrum from the recorded spectrum. The fourth panel of Fig.~\ref{fig:noiseSubtract} shows the spectrum obtained after subtracting the background noise spectrum. We see no obvious improvement, which may be attributed to a time-variant noise spectrum the cause of which is unidentified. %Why the noise sources are time-varying is a fact yet to be determined.

Beyond this naive approach, we apply a range of filters to the data.
In signals processing, the ideal filter would be one that:
%\begin{itemize}
i) completely attenuates the undesired parts of the spectrum, 
ii) does not change the rest of the spectrum, and 
iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
%\end{itemize}
However, these three conditions are cannot all hold at once and any filter must compromise between them.

@@

\subsubsection{Cascade notch filter}
\begin{figure*}
\begin{center}
\includegraphics[width=.9\textwidth]{figures/notch_filter_response}
\end{center}
\caption{\label{fig:notchMagResponse}
Magnitude response for the first five notches of the $15$ cascaded notch filter described by Eqns.~\ref{eqn:notch} and ~\ref{eqn:notch15}. 
}
\end{figure*}

@@

We use the built-in MATLAB filter design toolbox in this work. 
The amplitude response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fourth row of Fig.~\ref{fig:NotchWienerCombined}.
We see that the mains harmonics are significantly attenuated in comparison to the results shown in Fig.~\ref{fig:noiseSubtract}.

@@

\begin{figure*}
	\begin{center}
	\includegraphics[width=\textwidth]{figures/notch_and_wiener_superplot.pdf}
	\end{center}
	\caption{\label{fig:NotchWienerCombined}
	Time series (left column) and frequency spectrum (right column) results for the notch and Wiener filters.
	The first row shows the original speech recording.
	The second row shows the raw output from the optical microphone.
	The third row shows the naive technique of subtracting off the noise spectrum from the recorded spectrum.
	The fourth, fifth, and sixth rows show the results of applying the notch filter, the Wiener filter, and both combined, respectively.
	}
\end{figure*}

@@

In this work, we construct a higher-order causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}. A higher order Wiener filter provides greater smoothing of the input signal but also increases the computational memory required. For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency. The time series and frequency spectrum after applying the Wiener filter are shown in the fifth row of Fig.~\ref{fig:NotchWienerCombined}.
We see a significant improvement in the time series of the recovered signal, however a strong noise hum persists, audibly. 
%The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
%It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
%\han{should we add more details here too?} \jam{[seconded]}
%The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
%The timeseries is visibly cleaner, however there is a strong noise hum.

@@

By combining the notch and Wiener filter, we can trade off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:NotchWienerCombined}.
The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone. The mains noise is all but removed and more voice information is retained. However, the recovered voice still sounds muffled and is not understandable.

@@

\subsubsection{logMMSE estimator}
\label{sec:logmmse}
% logMMSE

>>>>

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/setup_pic2.pdf}
\caption{Photograph of the optical microphone. The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right. The interferometer is laid out as in Figure~\ref{fig:ifo_schematic_podo}. In the main image, the photodiode is placed behind the cloth screen. The inset shows a face-on view of the photodiode with the covering screen removed. See Appendix~\ref{app:circuit_diagram} for a circuit diagram and detailed photograph of the circuit.}
	\label{fig:setup_pic2}
\end{figure}

@@

\begin{figure*}
	\begin{center}
    %% trim order is <left> <lower> <right> <upper>
	\includegraphics[width=\textwidth,trim={1cm 4.5cm 1cm 3.5cm},clip]{figures/notch_and_wiener_superplot_v2.pdf}
	\end{center}
	\caption{\label{fig:BackgroundNotchWienerCombined}
	Time series (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
	The first and second rows show the original speech recording and raw output from the optical microphone respectively. 
    The third row shows a background noise recording from the optical microphone when no audio signal is played. 
    The fourth row shows the result of subtracting the background noise spectrum in the third row from the recording in the second row. 
	%The second row shows the raw output from the optical microphone.
	%The third row shows the naive technique of subtracting off the noise spectrum from the recorded spectrum.
	The fifth, sixth, and seventh rows show the results of applying the notch filter, the Wiener filter, and both combined, respectively.
	}
\end{figure*}

%\begin{figure*}
%\begin{center}
%% trim order is <left> <lower> <right> <upper>
%\includegraphics[width=\textwidth,trim={1cm 8.5cm 1cm 8cm},clip]{figures/freqSpectrumOriginalNoiseSubtracted.pdf}
%\end{center}
%\caption{\label{fig:noiseSubtract}
%Frequency spectrum of a $1\,{\rm s}$ signal before the use of filters. 
%The first panel shows the original frequency spectrum of the human voice recording (saying ``A cathode ...''). 
%The second panel shows the spectrum recorded after the signal passes through the optical microphone (sampled at $16\,000\,{\rm Hz}$).
%The third panel shows a noise-only recording from the optical microphone when no audio signal is being played. 
%The fourth panel shows the result of subtracting the noise-only spectrum in the third panel from the recording in the second panel. 
%We see no obvious improvement from this method. 
%}
%\end{figure*}

Results for a selection of filters are collated in Fig.~\ref{fig:BackgroundNotchWienerCombined} where the left and right columns show timeseries and Fourier spectrum results respectively. 
The first and second rows of Fig.~\ref{fig:BackgroundNotchWienerCombined} shows the source signal and the raw optical microphone recording respectively. 
The third row shows a background spectrum of the optical microphone, as shown previously in the top panel Fig.~\ref{fig:psd_noise}.

%The Fourier spectrum of the source signal is shown in the first panel of Fig.~\ref{fig:BackgroundNotchWienerCombined}. The second panel shows the spectrum of the raw optical microphone recording of the signal. The third panel of Fig.~\ref{fig:noiseSubtract} shows a background spectrum of the optical microphone, as shown previously in the top panel Fig.~\ref{fig:psd_noise}.

Given that we have access to the background noise spectrum, an intuitive way to remove noise is via subtracting the noise spectrum from the recorded spectrum. 
The fourth panel of Fig.~\ref{fig:BackgroundNotchWienerCombined} shows the spectrum obtained after subtracting the background noise spectrum. 
We see no obvious improvement, which may be attributed to a time-variant noise spectrum the cause of which is unidentified. %Why the noise sources are time-varying is a fact yet to be determined.

Beyond this na\"{i}ve approach, we apply a range of filters to the data.
In signals processing, the ideal filter would be one that:
%\begin{itemize}
i) completely attenuates the undesired parts of the spectrum, 
ii) does not change the rest of the spectrum, and 
iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
%\end{itemize}
However, these three conditions are cannot all hold at once and any filter must compromise between them.

@@

\subsubsection{Cascade notch filter}
\begin{figure*}
\begin{center}
\includegraphics[width=.9\textwidth]{figures/notch_filter_response.jpg}
\end{center}
\caption{\label{fig:notchMagResponse}
Magnitude response for the first five notches of the $15$ cascaded notch filter described by Eqns.~\ref{eqn:notch} and ~\ref{eqn:notch15}. 
}
\end{figure*}

@@

We use the built-in MATLAB filter design toolbox in this work. 
The amplitude response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
We see that the mains harmonics are significantly attenuated in comparison background subtracted results shown in the fourth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.

@@

@@

In this work, we construct a higher-order causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}. A higher order Wiener filter provides greater smoothing of the input signal but also increases the computational memory required. For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency. The time series and frequency spectrum after applying the Wiener filter are shown in the sixth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
We see a significant improvement in the time series of the recovered signal, however a strong noise hum persists, audibly. 
%The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
%It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
%\han{should we add more details here too?} \jam{[seconded]}
%The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
%The timeseries is visibly cleaner, however there is a strong noise hum.

@@

By combining the notch and Wiener filter, we can trade off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:BackgroundNotchWienerCombined}..
The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone. The mains noise is all but removed and more voice information is retained. However, the recovered voice still sounds muffled and is not understandable.

@@

\subsubsection{logMMSE estimator}
\label{sec:logmmse}
